Scalar Items in Embedded Position: {A}n Experimental Revisit

### 2012-08-28 text/paper/paper1.tex
changes in paper1

A number of empirical studies have already examined the availability
of readings for \as- and \es-sentences
\citep[e.g.][]{GeurtsPouscoulous2009:Embedded-Implic,CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}. However,
results have not been as clear-cut as one might have hoped for: for
instance, the available empirical evidence is inconclusive as to
whether local readings exist \citep[see also][for related
discussion]{Tielvan-Tiel2012:Embedded-Scalar}, and previous studies 
did not explicitly address all preference relations between readings relevant to
distinguish between theoretical positions. In the following, we will
argue that these heterogeneous results were partly due to the use of 
standard procedures such as picture-verification paradigm, which might not 
be sufficiently sensitive for the phenomenon under investigation,
for instance, due to possible confounding effects of the pictures used.
% We hypothesized that previous studies might be
% insufficiently informative because of the focus on (variants of) a
% picture-verification paradigm.\dn{relate to Bob van Tiel's work} The
% problem is that in order to test the availability of different
% candidate meanings different pictures have to be presented, so that
% effects of pictorial complexity or stereotypicality could never be
% ruled out entirely.
Moreover, previous studies have only accumulated limited evidence
pertaining to the second question that may help decide between
theoretical positions, namely which of the attested readings subjects
prefer. Finally, previous studies presented target sentences
visually. But as it is often argued that focal stress on an
embedded scalar item can favor a local reading
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,ChemlaSpector2010:Experimental-Ev,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar},
it seems important to present target sentences auditorily, so as to be
able to control for effects of ``silent prosody'' that subjects may
apply when reading a sentence see Bader 1998, \citep{Fodor1998:Learning}
\dn{provide
  references}.

@@

Finally, the above-mentioned studies did not control for possible
conflating effects of ``silent prosody''. A growing number of evidence
suggests that numerous factors might influence accent placement and
prosodic phrasing even while reading, amongst them default
accentuation, constituent length, rhythmic phenomena, and individual
variation (Augurzky, 2008; Bader, 1998; Fodor, 2000, 2001; Kentner,
2012, Steinhauer et.  al., 2001, see also the discussion in Frazier,
2009).\dn{references} As it has been suggested in the theoretical
literature that the availability of a local reading might hinge on the
realization of contrastive stress on the scalar item,
%Even traditionalism maintains that local
%readings are attested, but requiring a contrastive stress on the
%scalar item
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar}, it is necessary to
control for potential variation due to potential differences in silent accent placement.
We therefore presented sentences auditorily and addressed the frequently proposed claim 
that local readings are triggered by contrastive
stress.\dn{relate to \citet{SchwarzClifton2008:Strengthening-o} and
  others?}

@@

\end{exe}
In late-closure sentences, a modifier can be attached to one of two
preceding hosts. More specifically, in (\ref{bsp:target-related}), the
prepositional phrase \emph{with suns} can be attached to the preceding
noun \emph{squares}, resulting in the so-called \emph{late-closure} or
\emph{low-attachment} reading ((\ref{bsp:target-related-LC}), see
Frazier, 1987, for details)\dn{references}. Alternatively, the
\acro{pp} can be attached to the whole conjoined \acro{np}
\emph{circles and squares}, corresponding to the \emph{early-closure}
or \emph{high-attachment} reading
(\ref{bsp:target-related-EC}). Generally, the late-closure reading is
preferred over the early closure reading for sentences comparable to
those in (\ref{bsp:target-related}) (REFS)\dn{references}.

For each sentence, two sequences were designed. In the first, the
critical position associated with the dispreferred early-closure
reading preceded the critical position for the preferred late-closure
reading. In this case, a partly covered picture as in
Figure~\ref{fig:exec1} preceded a picture as in \ref{fig:exec2}.
%
\begin{figure}[ht]
	\centering
	\subfloat[][Step 1]{ 
		\fbox{\includegraphics[width=5.5cm]{../pictures/paper/ec_01_3.pdf}}
	    \label{fig:exec1}
	}
	\subfloat[][Step 2]{
		\fbox{\includegraphics[width=5.5cm]{../pictures/paper/ec_01_5.pdf}}
	    \label{fig:exec2}
	}
	\caption[]{Critical steps of a sequence for sentence
          (\ref{bsp:target-related}) where the early-closure reading
          (\ref{bsp:target-related-EC}) can be judged first.}
	\label{fig:exec}
\end{figure}
%
If participants adopted an early-closure reading as in
(\ref{bsp:target-related-EC}), a \emph{false}-judgment would be
expected at the position illustrated in \ref{fig:exec1}, as none of
the circles connected to the letter contain any suns. Under the
late-closure reading in (\ref{bsp:target-related-EC}), in contrast, a
\emph{true}-response would be expected no sooner than in the step
shown in Figure~\ref{fig:exec2}. The order of possible judgements is
reversed in the second kind of sequence including pictures like
\ref{fig:exlc1} followed by pictures like \ref{fig:exlc2}. Here, the
late-closure reading in (\ref{bsp:target-related-LC}) can be judged
first requiring a \emph{true}-response on \ref{fig:exlc1}. The
early-closure reading in (\ref{bsp:target-related-EC}) can only be
judged later in the sequence. When we reach \ref{fig:exlc2} a
\emph{false}-response is expected under an
early-closure/high-attachment reading.

@@

\paragraph{Testing Effects of prosody.}\dn{These two paragraphs
  still need a proper home; Petra opts for ``materials'' section.}

The availability of local readings has often been claimed to be
affected by the presence or absence of a focal accent on the scalar
item. By contrastively accenting the scalar, potential alternatives
are highlighted, resulting in more implicatures being computed.
Some approaches have even claimed that local readings exclusively 
emerge when the scalar item is accented and are otherwise generally
unavailable (Refs).

Though there seems to be general consensus about these potential 
prosodic effects, empirical evidence for this assumption 
is still scarce. One notable exception comes from a pilot study 
by Frazier, in which participants were instructed to read embedded
and non-embedded versions of English ae sentences, in which some was either
capitalized or non-capitalized. Prior to the experiment, participants 
were told that capitalizing corresponded to contrastive stress.
Interestingly, strengthened interpretations occurred rather often 
in the embedded sentences in this task (59 \%), but were 
not affected by capitalizing. ...
Another study relevant for the present consideration comes from Schwarz et al., 
examining the effects of contrastive accentuation of "or" in non-embedded sentences 
like {\it Mary will invite Fred or/OR Sam to barbecue}. After listening to these 
sentences, participants were presented two alternative readings ({\it 1. She will 
invite Fred or Sam or possibly both}, or {\it 2. She will invite Fred or Sam but 
not both}).

%  It
%has been claimed that local readings only emerge in the case of
%special accentuation of the the scalar item under consideration
%(REFERENCES). Further, it is generally assumed that the required type
%of accentuation is a pitch accent (REFS). To test for this possibility
%we decided to present the sentences (eg., (\ref{ex:as}) and
%(\ref{ex:es})) auditorily in two versions each. In the accented
%version a pitch accent was placed on {\it einige}. The second version
%had neutral prosody. If accentuation is the driving force for local
%readings, we would expect a higher proportion of local readings in the
%former than in the latter version of the sentences.

We also wished to make sure that differences in prosody may generally
lead to a shift in interpretation preferences in the \acro{itv}
task. Sentences like (\ref{bsp:target-related}) are well suited to
test this assumption since it is well-known that their
interpretational preferences are sensitive to prosody. If an utterance
of (\ref{bsp:target-related}) includes a prosodic phrase boundary
between {\it squares} and {\it with}, this intonational pattern
strongly corresponds to the high-attachment reading. If, instead, a
prosodic phrase boundary after {\it circles} is present, the
acceptance of the low-attachment reading is boosted. To test whether
prosodic information is taken into account in the \acro{itv} we
presented both types of prosodic structures expecting the described
boost in acceptance for the low-attachment reading. In addition, we
also included a neutral version containing no phrase marking
whatsoever. This should receive intermediate acceptance
rates. Further, we also presented unambiguous sentences corresponding
to the early-closure reading. As in the case of the \as- and
\es-sentences the unambiguous counterparts served as a baseline to
control for response biases that are independent of sentence meaning.

@@

We constructed a set of 15 items for \as- and \es-sentences,
respectively, analogous to the examples in \ref{ex:as} and
\ref{ex:es}. Sentences were introduced by a subject {\small
  \acro{DP}}, including the quantifiers {\it alle diese} ("all of
these") or {\it genau eine(r) der} ("exactly one of these"), as well
as the head noun which denoted different icons, such as letters or
bells. Subjects were followed by an auxiliary, and a {\small
  \acro{PP}} containing {\it einige}, a possessive pronoun, and a noun
denoting a geometrical object, e.g. {\it einigen seiner
  Dreiecke/Quadrate} ("some of its triangles/squares"). The possessive
pronoun was used in order to fix relative quantifier scope to surface
scope, thus ensuring that {\it einige} has embedded status. The last
word was the main verb {\it verbunden} ("connected").  Each sentence
was recorded in a stressed and an unstressed version of scalar {\it
  einige} (see section xx below). As described above, three control
sentences were created for each experimental item, corresponding to
the critical positions in the course of uncovering the accompanying
sequences (see below).  The positional controls were recorded with
neutral prosody.  Items were evenly distributed across 5 lists using a
Latin square design.

@@

\paragraph{Acoustic properties}

A phonetically trained female native speaker of German was instructed
to realize two prosodic versions of each target sentence, and three
versions of the ambiguous preference-related controls. In addition to these
sentences, the unambiguous preference-related controls, the positional controls
and a set of unrelated fillers were also recorded. For these constructions, the speaker was instructed
to realize prosodic contours as neutral and natural as possible.

For all target sentences (\as and \es), the determiner \emph{einige} was
produced with a contrastive pitch accent as well as with a neutral
accent. Contrastive accents in German are realized by an
H*L contour (XYZ)\dn{insert references}, thus employing a generally
falling pattern. A set of 15 experimental items was recorded for each
condition (\as vs.~\es, \emph{accented} vs.~\emph{unaccented}), resulting in a
total number of 60 target sentences.

In contrast to the accent manipulation, the ambiguous preference-related controls
differed with respect to prosodic phrasing. Prosodic phrase boundaries
in German are realized by a rise in F0 as well as by a durational
increase on the final part of the constituent preceding the boundary
(prefinal lengthening) plus an optional pause (XYZ)\dn{references:
  Vaissiere, 1983; Fery, 1993}. Boundaries for these control sentences
were either realized at the position separating the second {\small \acro{PP}}
from the preceding material (\emph{late boundary}, corresponding to a
high-attachment reading) or directly preceding the second conjunct in
the first {\small \acro{PP}} (\emph{early boundary}, corresponding to a
low-attachment reading). As the prosodic realization of the targets
involved the comparison between an accented and a neutral variant, we
also included a third version of the ambiguous preference related controls, in which our
speaker produced the sentences without any pronounced
boundaries. For each prosodic variant, a set of
30 items was read, yielding 90 preference-related fillers.

@@

Before entering the judgment task, experimental items and
preference-related fillers were analyzed with respect to their acoustic
properties. As both accented elements as well as prosodic boundaries
were expected to differ with respect to their F0 and/or durational
properties, we calculated durational values as well as difference
values between minimal and maximal F0 for each word. As targets
slightly differed with respect to the total number of words as well as
with respect to certain lexical properties (i.e., \emph{seinen}
vs.~\emph{ihren}, see above (XYZ)\dn{proper reference}), we considered
the following analysis regions:

\begin{exe}
  \ex
    \begin{xlist}
      \ex $|_{\text{R}1}$~Alle  	\ $|_{\text{R}2}$~dieser 
      \ $|_{\text{R}3}$~$\acro{np}_1$  
      \ $|_{\text{R}4}$~sind  \ $|_{\text{R}5}$~mit  \
      $|_{\text{R}6}$~einigen  \ $|_{\text{R}7}$~ihrer 
       \ $|_{\text{R}8}$~{$\acro{np}_2$}  \ $|_{\text{R}9}$~verbunden.
    \ex       $|_{\text{R}1}$~Genau einer  	\ $|_{\text{R}2}$~der 
      \ $|_{\text{R}3}$~$\acro{np}_1$  
      \ $|_{\text{R}4}$~ist  \ $|_{\text{R}5}$~mit  \
      $|_{\text{R}6}$~einigen  \ $|_{\text{R}7}$~seiner 
       \ $|_{\text{R}8}$~{$\acro{np}_2$}  \ $|_{\text{R}9}$~verbunden.
    \end{xlist}
    % \begin{tabbing}
    %   $|_{\text{R}1}$ Genau einer \= 	\ $|_{\text{R}2}$ diese \=
    %   \ $|_{\text{R}3}$ \acro{np}$_1$ \= 
    %   \ $|_{\text{R}4}$ sind \= \ $|_{\text{R}5}$ mit \= \
    %   $|_{\text{R}6}$ einigen \= \ $|_{\text{R}7}$ seiner 
    %   \= \ $|_{\text{R}8}$  \acro{np}$_2$ \= \ $|_{\text{R}9}$
    %   verbunden. \kill
    %   $|_{\text{R}1}$ Alle \> 	\ $|_{\text{R}2}$ dieser \>
    %   \ $|_{\text{R}3}$ \acro{np}$_1$ \> 
    %   \ $|_{\text{R}4}$ sind \> \ $|_{\text{R}5}$ mit \> \
    %   $|_{\text{R}6}$ einigen \> \ $|_{\text{R}7}$ ihrer 
    %   \> \ $|_{\text{R}8}$  \acro{np}$_1$ \> \ $|_{\text{R}9}$
    %   verbunden. \\
    %   $|_{\text{R}1}$ Genau einer \> 	\ $|_{\text{R}2}$ der \>
    %   \ $|_{\text{R}3}$ \acro{np}$_1$ \> 
    %   \ $|_{\text{R}4}$ ist \> \ $|_{\text{R}5}$ mit \> \
    %   $|_{\text{R}6}$ einigen \> \ $|_{\text{R}7}$ seiner 
    %   \> \ $|_{\text{R}8}$  \acro{np}$_1$ \> \ $|_{\text{R}9}$ verbunden.
    % \end{tabbing}
\end{exe}

@@

For the durational analyses, constituents were automatically labeled
by the \emph{Aligner} tool (XYZ),\dn{reference! Rapp, 1998} and the
obtained values were manually corrected afterwards. For the targets,
two-factorial \acros{anova} with the factors \acro{Quantifier} (all
vs. exactly one) and \acro{Prosody} (accented vs.~unaccented) were
carried out. For the preference-related fillers, we carried out
one-factorial \acros{anova} with the factor \acro{Prosody} (early
boundary, late boundary, neutral prosody). F0 values were extracted by
means of special Praat scripts
(\url{http://www.fon.hum.uva.nl/praat/}). For the present analyses,
differences between minimal and maximal F0 values for each region or
word were calculated. Again, two-factorial \acros{anova}s were carried
out for statistical comparison of the target sentences, and one-factorial
\acros{anova}s were carried out for preference-related controls.

The details of these analyses are found in
Appendix~\ref{sec:audit-sent-mater}. In sum, our speaker reliably
produced (i) differences in accent realization for the target
sentences and (ii) the expected boundary realizations for the
preference-related controls. Whereas accented elements clearly
differed from their unaccented counterpart by showing an increase in
duration and F0 range, prosodic boundaries for our preference-related
fillers were realized by pre-final lengthening (i.e. an increase in F0
and duration at the position preceding the boundary).

>>>>

A number of empirical studies have already examined the availability
of readings for \as- and \es-sentences
\citep[e.g.][]{GeurtsPouscoulous2009:Embedded-Implic,CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}. However,
results have not been as clear-cut as one might have hoped for: for
instance, the available empirical evidence is inconclusive as to
whether local readings exist \citep[see also][for related
discussion]{Tielvan-Tiel2012:Embedded-Scalar}, and previous studies 
did not explicitly address all preference relations between readings relevant to
distinguish between theoretical positions. In the following, we will
argue that these heterogeneous results were partly due to the use of 
standard procedures such as picture-verification paradigm, which might not 
be sufficiently sensitive for the phenomenon under investigation,
for instance, due to possible confounding effects of the pictures used.
% We hypothesized that previous studies might be
% insufficiently informative because of the focus on (variants of) a
% picture-verification paradigm.\dn{relate to Bob van Tiel's work} The
% problem is that in order to test the availability of different
% candidate meanings different pictures have to be presented, so that
% effects of pictorial complexity or stereotypicality could never be
% ruled out entirely.
Moreover, previous studies have only accumulated limited evidence
pertaining to the second question that may help decide between
theoretical positions, namely which of the attested readings subjects
prefer. Finally, previous studies presented target sentences
visually. But as it is often argued that focal stress on an
embedded scalar item can favor a local reading
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,ChemlaSpector2010:Experimental-Ev,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar},
it seems important to present target sentences auditorily, so as to be
able to control for effects of ``silent prosody'' that subjects may
apply when reading a sentence see Bader 1998, \citep{Fodor98}
\dn{provide
  references}.

@@

Finally, the above-mentioned studies did not control for possible
conflating effects of ``silent prosody''. A growing number of evidence
suggests that numerous factors might influence accent placement and
prosodic phrasing even while reading, amongst them default
accentuation, constituent length, rhythmic phenomena, and individual
variation (Augurzky, 2008; Bader, 1998; Fodor98, 2000, 2001; Kentner,
2012, Steinhauer et.  al., 2001, see also the discussion in Frazier,
2009).\dn{references} As it has been suggested in the theoretical
literature that the availability of a local reading might hinge on the
realization of contrastive stress on the scalar item,
%Even traditionalism maintains that local
%readings are attested, but requiring a contrastive stress on the
%scalar item
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar}, it is necessary to
control for potential variation due to potential differences in silent accent placement.
We therefore presented sentences auditorily and addressed the frequently proposed claim 
that local readings are triggered by contrastive
stress.\dn{relate to \citet{SchwarzClifton2008:Strengthening-o} and
  others?}

@@

\end{exe}
In late-closure sentences, a modifier such as a relative clause or a 
prepositional phrase can be attached to one of two
preceding hosts. For instance, in (\ref{bsp:target-related}), the
prepositional phrase \emph{with suns} can be attached to the preceding
noun \emph{squares}, resulting in the so-called \emph{late-closure} or
\emph{low-attachment} reading ((\ref{bsp:target-related-LC}). Alternatively, 
the \acro{pp} can be attached to the whole conjoined \acro{np}
\emph{circles and squares}, corresponding to the \emph{early-closure}
or \emph{high-attachment} reading (\ref{bsp:target-related-EC}).
Late-closure readings have been observed to be preferred over the early 
closure reading for sentences comparable to those in (\ref{bsp:target-related})
\citep[e.g.][]{Frazier87,Frazier82}.

For each sentence, two sequences were designed. In the first one, the
critical position associated with the dispreferred early-closure
reading preceded the critical position for the preferred late-closure
reading. In this case, a partly covered picture as in
Figure~\ref{fig:exec1} preceded a picture as in \ref{fig:exec2}.
%
\begin{figure}[ht]
	\centering
	\subfloat[][Step 1]{ 
		\fbox{\includegraphics[width=5.5cm]{../pictures/paper/ec_01_3.pdf}}
	    \label{fig:exec1}
	}
	\subfloat[][Step 2]{
		\fbox{\includegraphics[width=5.5cm]{../pictures/paper/ec_01_5.pdf}}
	    \label{fig:exec2}
	}
	\caption[]{Critical steps of a sequence for sentence
          (\ref{bsp:target-related}) where the early-closure reading
          (\ref{bsp:target-related-EC}) can be judged first.}
	\label{fig:exec}
\end{figure}
%
If participants adopted an early-closure reading as in
(\ref{bsp:target-related-EC}), a \emph{false}-judgment would be
expected at the position illustrated in \ref{fig:exec1}, as none of
the circles connected to the letter contain any suns. Under the
late-closure reading in (\ref{bsp:target-related-EC}), in contrast, a
\emph{true}-response would be expected no sooner than in the step
shown in Figure~\ref{fig:exec2}. The order of possible judgements is
reversed in the second kind of sequence including pictures like
\ref{fig:exlc1} followed by pictures like \ref{fig:exlc2}. Here, the
late-closure reading in (\ref{bsp:target-related-LC}) can be judged
first requiring a \emph{true}-response on \ref{fig:exlc1}. The
early-closure reading in (\ref{bsp:target-related-EC}) can only be
judged later in the sequence. When we reach \ref{fig:exlc2} a
\emph{false}-response is expected under an
early-closure/high-attachment reading.

@@

\paragraph{Testing Effects of prosody.}\dn{These two paragraphs
  still need a proper home; Petra opts for ``materials'' section.
  - Petra's comment: wenn wir's wie jetzt etwas ausführlicher diskutieren, 
  kann's auch gerne hier bleiben}

The availability of local readings has often been claimed to be
affected by the presence or absence of a focal accent on the scalar
item. By contrastively accenting the scalar, potential alternatives
are highlighted, resulting in more implicatures being computed. Some 
approaches have even claimed that local readings exclusively 
emerge when the scalar item is accented and are otherwise generally
unavailable \citep[e.g.][]{Horn2006:The-Border-Wars,
Geurts2009:Scalar-Implicat,Geurts2010:Quantity-Implic,
Tielvan-Tiel2012:Embedded-Scalar}.

Though there seems to be general consensus about these potential 
prosodic effects, empirical evidence for this assumption 
is still scarce. One notable exception comes from a pilot study 
reported in \citet{Frazier08}, in which participants were 
instructed to read embedded and non-embedded versions of English 
\as-sentences. In the embedded sentences (e.g. {\it All of the students 
wrote some of the official memos}), {\it some} was either capitalized 
or non-capitalized. Prior to the experiment, 
participants were informed that capitalizing corresponded to contrastive 
stress. The task consisted in a forced-choice questionnaire with two
given alternatives (such as {\it All of the students wrote some but 
not all of the official memos}  vs. {\it All of the students wrote at 
least some of the official memos}). Interestingly, strengthened  
interpretations occurred rather often across all sentences in this task 
(59 \%), but were not affected by capitalizing. \fn{Note that, 
unfortunately, the authors do not provide percentages for the embedded 
conditions alone, but only the means across sentence types.} The authors 
conclude that accentuation did not affect their silent reading data, 
but raise the possibility that prosodic effects may have been present 
at least in a subset of the items (see footnote 1 in \citet{Frazier08}, 
p. 330).

Another study relevant for the present considerations comes from 
\citet{SchwarzClifton2008:Strengthening-o}, examining the effects of 
contrastive accentuation of "or" in non-embedded sentences like {\it 
Mary will invite Fred or/OR Sam to barbecue}. After listening to these 
sentences, participants were presented two alternative readings 
({\it 1. She will invite Fred or Sam or possibly both}, or {\it 2. 
She will invite Fred or Sam but not both}). Contratry to the silent-prosodic 
effects reported in the pilot experiment by Frazier and colleagues, the 
authors found a significant impact of overtly realized contrastive accents 
on implicature processing: Besides a general advantage for non-strengthened 
readings, focal accents significantly increased the strengthened interpretation 
(from 16.4 \% to 28.6 \%). These results are compatible with the "focus 
strengthening hypothesis" proposed by the authors.

To sum up so far, though an effect of contrastive accentuation has been 
claimed tobe essential for triggering local implicatures, experimental evidence 
is currently mixed, suggesting potential effects of presentation mode (visual vs. 
auditory), item-specific characteristics, as well as the specific construction 
under investigation (e.g. embedded cases vs. non-embedded cases such as in 
\citet{SchwarzClifton2008:Strengthening-o}).

Whereas contrastive accents are usually assumed to be realized by a special 
contour in English (i.e. L+H*, see \citet{Pierrehumbert90}), 
the exact phonological classification of contrast in German is still under 
debate \citep[see][]{Uhmann91,Fery93,Grabe98,Toepel06, Sudhoff10}. 
These approaches have in common that from an acoustic perceptive, prosodic 
prominence of contrastively accented constituents is realized by

\begin{enumerate}[1:]
\item a higher higher f0 maximum or a higher pitch range (difference 
between minimal and maximal f0 values), and
\item a longer duration of the accented element as opposed to its 
non-accented counterpart
\end{enumerate}

We therefore recorded sentences by a trained speaker familiar with the 
concept of contrastive focus and instructed her to produce contrast as 
described in 1 and 2. Both \as-sentences and \es-sentences
were recorded in two versions each. In the accented version, a pitch 
accent was placed on {\it einige}. The second version had neutral prosody. 
If accentuation is the driving force for local readings, we would expect 
a higher proportion of local readings in the former than in the latter 
version of the sentences.

Preference-related controls also differed prosodically. Modifier attachment 
ambiguities comparable to the examples in (\ref{bsp:target-related}) and 
(\ref{bsp:target-related2}) have been shown to be sensitive to differences 
in prosodic phrasing. Though there has been discussion as to whether 
speakers reliably \textit{produce} disambiguating cues in such constructions 
\citep[e.g.][]{Allbritton96,Kraljik05, Snedeker03, Schafer00}, overt prosodic 
boundaries are generally acknowledged to guide interpretive processes of Late 
Closure sentences in \textit{comprehension} \citep[e.g.][]{Steinhauer99, Augurzky06}. 
Specifically, it has been shown that prosodically separating a modifier from 
the directly preceding material supports an early closure reading, whereas 
separating the two potential attachment sites is supportive of a late closure 
reading. Thus, a phrase boundary between {\it squares} and {\it with} in
(\ref{bsp:target-related}) corresponds to a high-attachment reading, whereas
a prosodic phrase boundary following {\it circles} corresponds to a low-attachment
reading.

To test whether prosodic information is taken into account in the \acro{itv} task, we
presented sentences in both of these prosodic variants. In addition, in order to test 
prosody-independent preferences, we also included a neutral version without any 
pronounced boundary. Further, we also presented 
unambiguous sentences corresponding to the early-closure reading. As in the case 
of the \as- and \es-sentences, the unambiguous counterparts served as a baseline to
control for response biases that are independent of sentence meaning.

@@

We constructed a set of 15 items for \as- and \es-sentences,
respectively, analogous to the examples in \ref{ex:as} and
\ref{ex:es}. Sentences were introduced by a subject {\small
  \acro{DP}}, including the quantifiers {\it alle diese} ("all of
these") or {\it genau eine(r) der} ("exactly one of these"), as well
as the head noun which denoted different icons, such as letters or
bells. Subjects were followed by an auxiliary, and a {\small
  \acro{PP}} containing {\it einige}, a possessive pronoun, and a noun
denoting a geometrical object, e.g. {\it einigen seiner
  Dreiecke/Quadrate} ("some of its triangles/squares"). The possessive
pronoun was used in order to fix relative quantifier scope to surface
scope, thus ensuring that {\it einige} has embedded status. The last
word was the main verb {\it verbunden} ("connected").  Each sentence
was recorded in a stressed and an unstressed version of scalar {\it
  einige}. As described above, three control
sentences were created for each experimental item, corresponding to
the critical positions in the course of uncovering the accompanying
sequences.  The positional controls were recorded with
neutral prosody.  Items were evenly distributed across 5 lists using a
Latin square design.

@@

\paragraph{Acoustic properties}
A phonetically trained female native speaker of German was instructed
to realize two prosodic versions of each target sentence, and three
versions of the ambiguous preference-related controls. In addition to these
sentences, the unambiguous preference-related controls, the positional controls
and a set of unrelated fillers were also recorded. For these constructions, the speaker was instructed
to realize prosodic contours as neutral and natural as possible.

For all target sentences (\as and \es), the determiner \emph{einige} was
produced with a contrastive pitch accent as well as with a neutral
accent. Contrastive accents in German are realized by an increase in f0 (range)
as well as by a longer duration (see {sec:design} above). A set of 15 
experimental items was recorded for each condition (\as vs.~\es, \emph{accented} 
vs.~\emph{unaccented}), resulting in a total number of 60 target sentences.

In contrast to the accent manipulation, the ambiguous preference-related controls
differed with respect to prosodic phrasing. Prosodic phrase boundaries
in German are realized by a rise in F0 as well as by a durational
increase on the final part of the constituent preceding the boundary
(prefinal lengthening) plus an optional pause \citep[e.g.][]{Vaissiere83,Fery93}.
Boundaries for these control sentences were either realized at the position 
separating the second {\small \acro{PP}}
from the preceding material (\emph{late boundary}, corresponding to a
high-attachment reading) or directly preceding the second conjunct in
the first {\small \acro{PP}} (\emph{early boundary}, corresponding to a
low-attachment reading). As the prosodic realization of the targets
involved the comparison between an accented and a neutral variant, we
also included a third version of the ambiguous preference related controls, in which our
speaker produced the sentences without any pronounced
boundaries. For each prosodic variant, a set of
30 items was read, yielding 90 preference-related fillers.

@@

Before entering the judgment task, experimental items and
preference-related fillers were analyzed with respect to their acoustic
properties. As both accented elements as well as prosodic boundaries
were expected to differ with respect to their F0 and/or durational
properties, we calculated durational values as well as difference
values between minimal and maximal F0 for each word. As targets
slightly differed with respect to the total number of words as well as
with respect to certain lexical properties (i.e., \emph{seinen}
vs.~\emph{ihren}, we considered the following analysis regions:

\begin{exe}
  \ex
    \begin{xlist}
      \ex $|_{\text{R}1}$~Alle  	\ $|_{\text{R}2}$~diese 
      \ $|_{\text{R}3}$~$\acro{np}_1$  
      \ $|_{\text{R}4}$~sind  \ $|_{\text{R}5}$~mit  \
      $|_{\text{R}6}$~einigen  \ $|_{\text{R}7}$~ihrer 
       \ $|_{\text{R}8}$~{$\acro{np}_2$}  \ $|_{\text{R}9}$~verbunden.
    \ex       $|_{\text{R}1}$~Genau einer  	\ $|_{\text{R}2}$~der 
      \ $|_{\text{R}3}$~$\acro{np}_1$  
      \ $|_{\text{R}4}$~ist  \ $|_{\text{R}5}$~mit  \
      $|_{\text{R}6}$~einigen  \ $|_{\text{R}7}$~seiner 
       \ $|_{\text{R}8}$~{$\acro{np}_2$}  \ $|_{\text{R}9}$~verbunden.
    \end{xlist}
    % \begin{tabbing}
    %   $|_{\text{R}1}$ Genau einer \= 	\ $|_{\text{R}2}$ diese \=
    %   \ $|_{\text{R}3}$ \acro{np}$_1$ \= 
    %   \ $|_{\text{R}4}$ sind \= \ $|_{\text{R}5}$ mit \= \
    %   $|_{\text{R}6}$ einigen \= \ $|_{\text{R}7}$ seiner 
    %   \= \ $|_{\text{R}8}$  \acro{np}$_2$ \= \ $|_{\text{R}9}$
    %   verbunden. \kill
    %   $|_{\text{R}1}$ Alle \> 	\ $|_{\text{R}2}$ dieser \>
    %   \ $|_{\text{R}3}$ \acro{np}$_1$ \> 
    %   \ $|_{\text{R}4}$ sind \> \ $|_{\text{R}5}$ mit \> \
    %   $|_{\text{R}6}$ einigen \> \ $|_{\text{R}7}$ ihrer 
    %   \> \ $|_{\text{R}8}$  \acro{np}$_1$ \> \ $|_{\text{R}9}$
    %   verbunden. \\
    %   $|_{\text{R}1}$ Genau einer \> 	\ $|_{\text{R}2}$ der \>
    %   \ $|_{\text{R}3}$ \acro{np}$_1$ \> 
    %   \ $|_{\text{R}4}$ ist \> \ $|_{\text{R}5}$ mit \> \
    %   $|_{\text{R}6}$ einigen \> \ $|_{\text{R}7}$ seiner 
    %   \> \ $|_{\text{R}8}$  \acro{np}$_1$ \> \ $|_{\text{R}9}$ verbunden.
    % \end{tabbing}
\end{exe}

@@

For the durational analyses, constituents were automatically labeled
by the \emph{Aligner} tool \citet{Rapp98},and the
obtained values were manually corrected afterwards. For the targets,
two-factorial \acros{anova} with the factors \acro{Quantifier} (all
vs. exactly one) and \acro{Prosody} (accented vs.~unaccented) were
carried out. For the preference-related fillers, we carried out
one-factorial \acros{anova} with the factor \acro{Prosody} (early
boundary, late boundary, neutral prosody). F0 values were extracted by
means of special Praat scripts
(\url{http://www.fon.hum.uva.nl/praat/}). For the present analyses,
differences between minimal and maximal F0 values for each region or
word were calculated. Again, two-factorial \acros{anova} were carried
out for statistical comparison of the target sentences, and one-factorial
\acros{anova} were carried out for preference-related controls.

The details of these analyses are found in
Appendix~\ref{sec:audit-sent-mater}. \dn{Petras Kommentar: Vielleicht sollten wir hier jeweils 1 Figure für akzentuiert vs. nicht und boundary vs. nicht reinbringen. Fabian meint zurecht, dass man nicht einfach die Ergebnisse nur in den Appendix packen kann. Es wären dann halt 4 Figures, da wir ja f0 und duration ansehen, und wir würden nur die experimentell interessanten veranschaulichen und die Statistik dann im Appendix lassen - was meint Ihr?}

In sum, our speaker reliably
produced (i) differences in accent realization for the target
sentences and (ii) the expected boundary realizations for the
preference-related controls. Whereas accented elements clearly
differed from their unaccented counterpart by showing an increase in
duration and F0 range, prosodic boundaries for our preference-related
fillers were realized by pre-final lengthening (i.e. an increase in F0
and duration at the position preceding the boundary).
